VLI - a library for large integer and polynomial arithemtic

### 2013-01-24 doc/paper_vli/vli2.tex
wip vli paper

\section{Overview of the GPU architecture}
\begin{itemize}
\item A GPU has multiple streaming multiprocessors (SM)
\item They execute instructions in a SIMT (single-instruction, multiple-thread).
\item A group of 32 threads (called ``warp'') execute the instruction stream in a lockstep fashion on multiple data, if the threads follow different execution paths all branches need to be executed serially.
\item SM switches between up to 64 warps to hide latencies. $\rightarrow$ To gain good performance thousands of threads are necessary.
\item NVIDIA CUDA resembles this hardware properties by threads that are organized in so called blocks, which consist of a few hundred threads. A thread block is always scheduled on a single SM, but it is possible to schedule multiple thread blocks on a single SM.
\item Memory hierarchy: large (several GB) slow global memory accessible from all threads; small (a few KB) very fast shared memory which is shared among all threads of a thread block; The registers of the SM are split among all threads hosted by the SM.
\item Special purpose memories: Texture memory: globally shared, read-only, cache optimized for 2D access
\end{itemize}

\section{Optimized inner product}
\paragraph{CPU}
\begin{itemize}
\item The hot spot of HTSE are inner products of large vectors with symbolic polynomials as components. The coefficients of the polynomials are large integers.
\item We continue to use the single large integer kernels
\item OpenMP parallelization: split vector into chunks, each thread does part of the inner product
\item No SIMD, since we had to use 32bit words and the number of operations for a multiplication increases by a factor TODO due to the manual handling of the carry.
\end{itemize}
\paragraph{GPU}
\begin{itemize}
\item Many small tasks of the same form. $\rightarrow$ well suited for GPUs.
\item {\bf Implementation details}
\item Several coefficients $a_{ijkl}$, $b_{i'j'k'l'}$ of the polynomials to be multiplied contribute to the same coefficient $c_{IJKL}$ of the resulting polynomial
\begin{equation}
    c_{IJKL} = \sum_{i,j,k,l} a_{ijkl} \cdot b_{I-i,J-j,K-k,L-l}
\end{equation}
\item The number of terms in this sum depends on the orders $I,J,K,L$ of the resulting coefficient $\rightarrow$ load inbalance
\item We group 32 coefficients of the result into a task and distribute them among the warps of threads (Fig.\ \ref{fig:GPU_load_balance}). {\color{red} TODO: Tim please elaborate}
\begin{figure}[t]
    \begin{center}
    \mbox{
        \includegraphics[scale=0.37, angle=-90]{coeffs.eps} 
        \includegraphics[scale=0.37, angle=-90]{warp.eps} 
    }
    \caption{Left: Number of contributions to each coefficient of the resulting polynomial (red line) sorted in reverse order and split into tasks. Right: Load balancing of the tasks over the warps.}
    \label{fig:GPU_load_balance}
    \end{center}
\end{figure}
\item Coalesced memory access?
\item Using NVIDIA parallel thread execution (PTX) assembly language to benefit from hardware carry bit propagation \verb|addc.cc| and fused multiply add with carry bit propagation \verb|madc.cc|.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\section{Optimized inner product}
\paragraph{CPU}
\begin{itemize}
\item The hot spot of HTSE are inner products of large vectors with symbolic polynomials as components. The coefficients of the polynomials are large integers.
\item We continue to use the single large integer kernels
\item OpenMP parallelization: split vector into chunks, each thread does part of the inner product
\item No SIMD, since we had to use 32bit words and the number of operations for a multiplication increases by a factor TODO due to the manual handling of the carry.
\end{itemize}
\paragraph{GPU}
\begin{itemize}
\item Many small tasks of the same form. $\rightarrow$ well suited for GPUs.
\item {\bf Implementation details}
\item Element-wise polynomial multiplication + reduction at the end.
\item Each thread calculates one coefficient $c_{IJKL}$ the product polynomial.
\item Several coefficients $a_{ijkl}$, $b_{i'j'k'l'}$ of the polynomials to be multiplied contribute to the same coefficient $c_{IJKL}$
\begin{equation}
    c_{IJKL} = \sum_{i,j,k,l} a_{ijkl} \cdot b_{I-i,J-j,K-k,L-l}
\end{equation}
\item The number of terms in this sum depends on the orders $I,J,K,L$ of the resulting coefficient $\rightarrow$ load inbalance
\item We group 32 coefficients of the result into a task and distribute them among the warps of threads (Fig.\ \ref{fig:GPU_load_balance}).
\begin{figure}[t]
    \begin{center}
    \mbox{
        \includegraphics[scale=0.37, angle=-90]{coeffs.eps} 
        \includegraphics[scale=0.37, angle=-90]{warp.eps} 
    }
    \caption{Left: Number of contributions to each coefficient of the resulting polynomial (red line) sorted in reverse order and split into tasks. Right: Load balancing of the tasks over the warps.}
    \label{fig:GPU_load_balance}
    \end{center}
\end{figure}
\item {\color{red} TODO: Tim please elaborate}
%\item Coalesced memory access?
\item Using NVIDIA parallel thread execution (PTX) assembly language to benefit from hardware carry bit propagation \verb|addc.cc| and fused multiply add with carry bit propagation \verb|madc.cc|.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
